Inferring the Recombinant Genealogy of a Pandemic Pathogen in Real Time

### 2025-09-11 latex/main.tex
Trim down a bit

\section*{SUMMARY}
% The dataset is freakin amazing, and will be useful for all sorts of stuff
% for years to come
Millions of SARS-CoV-2 whole genome sequences were collected as part of the 
response to the COVID-19 pandemic, generating a dataset of unprecedented
depth, richness and potential for informing future studies.
% Trees are a useful abstraction of the data
Estimated genealogies, describing the likely inheritance structure among
the sampled sequences and their putative ancestors, are fundamental to
% JK: Not happy with this specific phrase, but it's good to emphasise the point
% that data is not insight, and the sheer volume of data here requires
% special effort.
gleaning information from this ocean of data, 
and form the primary input to many downstream analyses.
A basic assumption of methods to infer such genealogies from viral 
genetic sequence data, however, is that recombination is negligible and 
the genealogy can be described as a tree.
% Recombination *is* important.
Recombinant lineages have risen to global prevalance in SARS-CoV-2,
and therefore such single-tree summaries of the data are necessarily
incomplete and potentially misleading.
% What do we do?
We present sc2ts, a method to infer reticulate genealogies for SARS-CoV-2
that automaticaly detects and fully incorporates recombination, 
and apply it to 2.48 million genomes.
We show that this structure captures a broad range of evolutionary
features documented in the literature and provides novel 
insights into the relationships between previously identified
recombinant lineages.
% We need to emphasise the *resource*
Together, the inference method, inferred genealogy and 
a suite of efficient and flexible open-source analysis tools, 
represent a powerful resource upon which future research 
can build.

@@

ancestral recombination graphs, population genetics, recombination, tree sequence, SARS-CoV-2

>>>>

%%% The summary (abstract) should consist of a single paragraph of 150 words or
%%% fewer.

\section*{SUMMARY}

Millions of SARS-CoV-2 genome sequences were collected in
response to the COVID-19 pandemic, forming a dataset of unprecedented
depth and richness.
Estimated genealogies are fundamental to
gleaning information from this ocean of data, 
and form the primary input to many downstream analyses.
A basic assumption of methods to infer genealogies from viral 
genetic sequence data is that recombination is negligible and 
the genealogy is a tree.
Recombinant lineages have risen to global prevalance in SARS-CoV-2,
and therefore such single-tree representations are necessarily
incomplete and potentially misleading.
We present sc2ts, a method to infer reticulate genealogies for SARS-CoV-2
that automaticaly detects and incorporates recombination, 
and apply it to 2.48 million genomes.
We show that the inferred genealogy captures a broad range of 
known evolutionary features and provides novel 
insights into the relationships between previously identified
recombinant lineages.
Together, the inference method, inferred genealogy and 
suite of open-source analysis tools, 
represent a powerful resource for future research.

@@

Ancestral recombination graphs, population genetics, recombination, tree sequence, SARS-CoV-2
